---
title: "Consumption of Chinook salmon by Steller sea lions along the northwest coast of Washington"
author: "Zoë Lewis and Benjamin Nelson"
date: "2024-08-10"
format: 
  html: 
    toc: true
    toc-depth: 2
    code-fold: true
    number-sections: true
    notebook: true
    theme:
      light: flatly
      dark: darkly
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description of Model

Biomass of Chinook salmon consumed by Steller sea lions between December 2020-August 2021 along the northwest coast of Washington.

First, we describe prey consumption models for Steller sea lions (SSL) adapted from @scordino2022. This model was developed to determine the biomass consumed by Steller sea lions between 2010-2013 along the northwest coast of Washington using hard parts identification of prey items within scat. Thus, we use many of the same parameters, however, abundance parameters changed based on updated counts of Steller sea lions in the region. Using this model, we first estimate the total biomass consumed by these Steller sea lions.

Second, we used new DNA metabarcoding diet data to estimate the seasonal proportion of Chinook within Steller sea lion diet. Given that the data are zero-inflated, proportion data, we use ZOIB modeling (described in both the html and the rmd file titled "ZOIB model", as well as in @nelson2021a) to determine seasonal proportions of Chinook consumed .

Third, we used the ZOIB modeling protocol to estimate the seasonal diet proportion of age-0 Chinook salmon consumed to determine the biomass and the number of age-0 Chinook salmon consumed by Steller sea lions along the northwest coast of Washington State. Then, we estimate the number of individual age-0 Chinook salmon consumed by Steller sea lions given previously used length distributions found in @chasco2017.

## Required packages

In this step, all necessary R packages are downloaded to produce the necessary outputs for the model.

```{r Load packages, warning=FALSE, message=FALSE, output = FALSE}

rm(list=ls()) 
gc() 

library(stringr)
library(readxl)
library(dplyr)
library(colorspace)
library(viridis)
library(zoib)
library(ggplot2)
library(lemon)
library(knitr)
library(tidyr)
library(patchwork)
library(EnvStats)
library(pander)
library('scales')
library(tibble)
```

# Steller sea lion consumption Model - Scordino et al. 2022

There are three steps to calculate the total biomass of prey consumed by Steller sea lions along the coast of Washington state, similar to the steps taken in @scordino2022. 

1)  Calculate individual Steller sea lion daily consumption

2)  Model abundance and determine abundance functions

3)  Calculate total biomass consumed using consumption and abundance

## Individual Steller sea lion daily consumption 

This section describes the distributions used to determine the daily consumption (in kilograms) of Steller sea lions based on their age, and sex. Parameters are derived from @winship2006; @winship2001, and the overall biomass model presented first in @scordino2022.

### % body weight consumed

Percent body weight consumed based on body weight estimates and sex are the same as those used in "Dietary niche overlap and prey consumption estimates for Steller sea lion and California sea lion in northwest Washington 2010-2013" @scordino2022. Percent body weight consumed determined using either normal distribution or uniform distributions.

```{r percent body weight consumed} 

#Female P 
FLarge <- rnorm(1,0.05,0.01)
FMed <- runif(1,0.06,0.07)
FMedSm <- runif(1,0.07,0.08)
FSmall <- runif(1,0.08,0.09)

#Male P 
MLarge <- runif(1,0.03,0.04)
MMed <- rnorm(1,0.04,0.01)
MMedSm <- runif(1,0.05,0.06)
MSmall <- runif(1,0.06,0.07)
MVSmall <- runif(1,0.08,0.09)
```

### Daily consumption 

Functions used to determine the daily consumption of prey (kg/day) by SSL is defined in @scordino2022: "Formulas for estimating the daily consumption of a randomly drawn juvenile female juvenile male, adult female, and adult male using the consumption of body weight parameters defined above and the uniform distribution of reported weight at age for each age and sex class as reported in @winship2001. Juvenile weights are for sea lions ages 1 to 5. Juvenile females use non-pregnant 5 year old weight as the maximum weight for the category. Adults use age classes of 5 up to maximum growth from @winship2001."

```{r individual daily consumption}

## Individual daily consumption juvenile female (I)
IndDailyConsumptionJuvF<- function() 
{
  JuvF <- runif(1,89,195)
  if(JuvF > 194){JuvF*FMed}
  if(JuvF > 152){JuvF*FMedSm}
  if(JuvF > 89){JuvF*FSmall}
}

## Individual daily consumption adult female (I)
IndDailyConsumptionAdF<- function() 
{
  AdFemale <- runif(1,195,305)
  if(AdFemale > 223){AdFemale*FLarge}
  if(AdFemale > 195){AdFemale*FMed}
}

## Individual daily consumption juvenile male (I)
IndDailyConsumptionJuvM<- function() 
{
  JuvM <- runif(1,125,286)
  if(JuvM > 182){JuvM*MSmall}
  if(JuvM > 125){JuvM*MVSmall}
}

## Individual daily consumption adult male (I)
IndDailyConsumptionAdM<- function() 
{
  AdMale <- runif(1,286,681)
  if(AdMale > 623){AdMale*MLarge}
  if(AdMale > 455){AdMale*MMed}
  if(AdMale > 310){AdMale*MMedSm}
  if(AdMale > 286) {AdMale*MSmall}
}
```


## Modeling abundance

Counts are from surveys conducted from 2020-2021 data counts only.

Count data was only used from days that included all local haulouts surveyed. Raw data for the counts are available from Scordino and Akmajian 2021. Abundance functions were modeled using a negative binomial distribution.

**Table 1**: Negative binomial parameters for modeling Steller sea lion abundance along the northwest coast of Washington State between December 2020-August 2021.

| **Season** | **Mean** | **Size** | **n** |
|:----------:|:--------:|:--------:|:-----:|
|   Winter   |  976.33  |  36.34   |   3   |
|   Spring   |  888.29  |  32.10   |   7   |
|   Summer   |  969.77  |  25.80   |  13   |

```{r abundance functions}
# Winter Steller sea lion abundance
WinAbund <- function()
  {
  rnbinom(n=1, mu=976.3333, size=36.34341)
}

# Spring Steller sea lion abundance
SprAbund <- function()
  {
  rnbinom(n=1, mu=888.2857, size=32.07886)
}

# Summer Steller sea lion abundance
SumAbund <- function()
{
  rnbinom(n=1, mu=969.7692, size=25.796)
}
```

### Count correction factor functions

Correction factors are the same as correction factors used in @scordino2022,

Methods from @scordino2022 describe seasonal count correction factors as follows:

"Correction factors from Olesiuk 2018 were used assuming that the correction factor for fall and spring have the same CF as caculated for winter. This assumption is supported by the findings of Whitlock et al. 2020 of very low attendance in May and by the end of August that were similar to Olesiuk's estimate for winter CF. Overall the correction factors from Olesiuk 2018 were similar to the estimates of Whitlock et al. 2020. These correction factors were for Steller sea lions. Olesiuk found that 36% of non-pups (SD=2.1%) were hauled out in fall, winter, and spring during the time window of 10:00 - 18:00 when our counts typically occurred. During summer Olesiuk found that 67.4% of non-pups (SD= 5.6%) were hauled out during survey time periods. Note that Olesiuk (2018) errored in reporting that 2.1% was the winter CV and 5.6% was the summer CV of the proportion of population hauled out; both should have read SD instead of CV (Personal communication Peter Olesiuk, 2021)."

```{r correction factors}
WinSprFallCF <- function()
{
  1/(rnorm(1,0.36,0.021))
}


SummerCF <- function()
{
  1/rnorm(1,0.674,0.056)
  }
```

### Seasonal Demographic Functions (Sex and age)

We assume that relative proportion of demographics has not changed from the initial estimates calculated for the model used in @scordino2022, as demographic counts were not performed for every survey in 2020-2021.

Methods from @scordino2022:

"Count data from Scordino and Akmajian (2021) were used to determine the proportion of the hauled out Steller sea lions that were adult male, adult female, and juvenile."

Note: In contrast to abundance count data, demographic counts were obtained from all days 2010-2013 that had demographic data, and did not need to have all haulouts surveyed within a day.

Seasonal demographic functions modeled with beta distributions for each season for observed adult female, adult male and juvenile proportions.

**Table 2**: Beta distribution parameters for modeling demographic proportions for adult male, adult female and juvenile Steller sea lions along the northwest coast of Washington State. Demographic proportions obtained from @scordino2022, 2010-2013 demographic counts.

| Season | Avg F | Sd F  | Avg J | Sd J  | Avg M | Sd M  | n   |
|--------|:------|:------|:------|:------|:------|:------|:----|
| Spring | 0.379 | 0.099 | 0.426 | 0.098 | 0.195 | 0.071 | 11  |
| Summer | 0.261 | 0.055 | 0.408 | 0.055 | 0.330 | 0.071 | 9   |
| Winter | 0.390 | 0.104 | 0.406 | 0.046 | 0.205 | 0.074 | 4   |

The proportions of individuals hauled out of each demographic group is best modeled using a beta distribution. To calculate the beta distribution we first calculated the shape 1 and shape 2 of the distribution for each demographic group by season."

```{r  distribution params}
demographics <- list(
  winter = list(
    juvenile = c(mean = 0.406, sd = 0.046),
    female = c(mean = 0.390, sd = 0.104),
    male = c(mean = 0.205, sd = 0.074)
  ),
  spring = list(
    juvenile = c(mean = 0.426, sd = 0.098),
    female = c(mean = 0.379, sd = 0.099),
    male = c(mean = 0.195, sd = 0.071)
  ),
  summer = list(
    juvenile = c(mean = 0.408, sd = 0.055),
    female = c(mean = 0.261, sd = 0.055),
    male = c(mean = 0.330, sd = 0.071)
  )
)

# Calculate beta shapes function
calc_beta_shapes <- function(mean, sd) {
  shape1 <- (mean^2 - mean^3 - mean * sd^2) / sd^2
  shape2 <- (mean - 2 * mean^2 + mean^3 - sd^2 + mean * sd^2) / sd^2
  return(c(shape1, shape2))
}

# General function to sample from Beta distribution for a given season and demographic
sample_beta_demo <- function(season, demo_group) {
  demo <- demographics[[season]][[demo_group]]
  shapes <- calc_beta_shapes(demo["mean"], demo["sd"])
  rbeta(n = 1, shape1 = shapes[1], shape2 = shapes[2])
}
```

```{r demographic functions}
WinJuvDemo <- function() sample_beta_demo("winter", "juvenile")
WinFemDemo <- function() sample_beta_demo("winter", "female")
WinMaleDemo <- function() sample_beta_demo("winter", "male")

SprJuvDemo <- function() sample_beta_demo("spring", "juvenile")
SprFemDemo <- function() sample_beta_demo("spring", "female")
SprMaleDemo <- function() sample_beta_demo("spring", "male")

SumJuvDemo <- function() sample_beta_demo("summer", "juvenile")
SumFemDemo <- function() sample_beta_demo("summer", "female")
SumMaleDemo <- function() sample_beta_demo("summer", "male")
```

### Sex of juveniles ratio

Given that Steller sea lion juveniles are not observably different based on sex, we used a known ratio of juvenile female to juvenile male sea lions from previously described in @scordino2022.

Sex ratios were determined from data from @wright2017, from cohorts 2001-2007, where survival rates occurred through year 5. Overall sex ratio of juveniles was then determined by averaging survivorship at age for ages 1-5 by each sex.

The proportion of Juvenile sex are as follows:

```{r juvenile sex ratios}
PropJuvF <- 0.544
PropJuvM <- 0.456
```

## Biomass calculation 

Using the above parameters and functions, we can now model the total biomass consumed by Steller sea lions along the northwest Washington coast from December 2020- August 2021.

### Biomass calculation function

```{r biomass calculation function}
calc_biomass <- function(abund_func, cf_func, demo_func, prop, ind_daily_cons_func) {
  replicate(2000, 
            abund_func() * cf_func() * demo_func() * prop * 90 * (ind_daily_cons_func())
  )
}

# Quantile function for biomass
get_biomass_quantiles <- function(biomass_data) {
  quantile(biomass_data, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
}
```

### Seasonal prey biomass consumed
::: panel-tabset
#### Winter

```{r winter prey biomass, cache=TRUE}

# Winter Biomass Calculations
tot_biomass_WinterJuvF <- calc_biomass(WinAbund, WinSprFallCF, WinJuvDemo, PropJuvF, IndDailyConsumptionJuvF)
tot_biomass_WinterJuvM <- calc_biomass(WinAbund, WinSprFallCF, WinJuvDemo, PropJuvM, IndDailyConsumptionJuvM)
tot_biomass_WinterAdF  <- calc_biomass(WinAbund, WinSprFallCF, WinFemDemo, 1, IndDailyConsumptionAdF)
tot_biomass_WinterAdM  <- calc_biomass(WinAbund, WinSprFallCF, WinMaleDemo, 1, IndDailyConsumptionAdM)

# Get quantiles for each demographic
Win.Juv.F_tot_biomass <- get_biomass_quantiles(tot_biomass_WinterJuvF)
Win.Juv.M_tot_biomass <- get_biomass_quantiles(tot_biomass_WinterJuvM)
Win.Ad.F_tot_biomass  <- get_biomass_quantiles(tot_biomass_WinterAdF)
Win.Ad.M_tot_biomass  <- get_biomass_quantiles(tot_biomass_WinterAdM)

# Combine quanitiles and format table for presentation in .html 
win.biomass.demographic <- cbind(Win.Juv.F_tot_biomass,Win.Juv.M_tot_biomass,Win.Ad.F_tot_biomass, Win.Ad.M_tot_biomass)

# Format table for presentation in .html 
colnames(win.biomass.demographic) <- c("Juvenile Female", "Juvenile Male", "Adult Female", "Adult Male")

kable(win.biomass.demographic, caption= "Winter biomass consumed by demographic (in kg)")

# Total biomass for the entire population in metric tons
Winter_tot_biomass <- (tot_biomass_WinterJuvF + tot_biomass_WinterJuvM + tot_biomass_WinterAdF + tot_biomass_WinterAdM) / 1000
Winter_tot_biomass_sum <- get_biomass_quantiles(Winter_tot_biomass)

kable(Winter_tot_biomass_sum, caption = "Total metric tons of prey consumed in winter")

```

#### Spring

```{r spring prey biomass, cache=TRUE}
# Spring Biomass Calculations
tot_biomass_SpringJuvF <- calc_biomass(SprAbund, WinSprFallCF, SprJuvDemo, PropJuvF, IndDailyConsumptionJuvF)
tot_biomass_SpringJuvM <- calc_biomass(SprAbund, WinSprFallCF, SprJuvDemo, PropJuvM, IndDailyConsumptionJuvM)
tot_biomass_SpringAdF  <- calc_biomass(SprAbund, WinSprFallCF, SprFemDemo, 1, IndDailyConsumptionAdF)
tot_biomass_SpringAdM  <- calc_biomass(SprAbund, WinSprFallCF, SprMaleDemo, 1, IndDailyConsumptionAdM)

# Get quantiles for each demographic in Spring
Spr.Juv.F_tot_biomass <- get_biomass_quantiles(tot_biomass_SpringJuvF)
Spr.Juv.M_tot_biomass <- get_biomass_quantiles(tot_biomass_SpringJuvM)
Spr.Ad.F_tot_biomass  <- get_biomass_quantiles(tot_biomass_SpringAdF)
Spr.Ad.M_tot_biomass  <- get_biomass_quantiles(tot_biomass_SpringAdM)

spr.biomass.demographic <- cbind(Spr.Juv.F_tot_biomass,Spr.Juv.M_tot_biomass,Spr.Ad.F_tot_biomass, Spr.Ad.M_tot_biomass)

colnames(spr.biomass.demographic) <- c("Juvenile Female", "Juvenile Male", "Adult Female", "Adult Male")
kable(spr.biomass.demographic, caption= "Spring biomass consumed by demographic (in kgs)")

# Total biomass for the entire population in Spring (metric tons)
Spring_tot_biomass <- (tot_biomass_SpringJuvF + tot_biomass_SpringJuvM + tot_biomass_SpringAdF + tot_biomass_SpringAdM) / 1000
Spring_tot_biomass_sum <- get_biomass_quantiles(Spring_tot_biomass)

kable(Spring_tot_biomass_sum, caption = "Total metric tons of prey consumed in spring")
```

#### Summer

```{r summer prey biomass, cache=TRUE}
# Summer Biomass Calculations
tot_biomass_SummerJuvF <- calc_biomass(SumAbund, SummerCF, SumJuvDemo, PropJuvF, IndDailyConsumptionJuvF)
tot_biomass_SummerJuvM <- calc_biomass(SumAbund, SummerCF, SumJuvDemo, PropJuvM, IndDailyConsumptionJuvM)
tot_biomass_SummerAdF  <- calc_biomass(SumAbund, SummerCF, SumFemDemo, 1, IndDailyConsumptionAdF)
tot_biomass_SummerAdM  <- calc_biomass(SumAbund, SummerCF, SumMaleDemo, 1, IndDailyConsumptionAdM)

# Get quantiles for each demographic in Summer
Sum.Juv.F_tot_biomass <- get_biomass_quantiles(tot_biomass_SummerJuvF)
Sum.Juv.M_tot_biomass <- get_biomass_quantiles(tot_biomass_SummerJuvM)
Sum.Ad.F_tot_biomass  <- get_biomass_quantiles(tot_biomass_SummerAdF)
Sum.Ad.M_tot_biomass  <- get_biomass_quantiles(tot_biomass_SummerAdM)

sum.biomass.demographic <- cbind(Sum.Juv.F_tot_biomass,Sum.Juv.M_tot_biomass,Sum.Ad.F_tot_biomass, Sum.Ad.M_tot_biomass)

colnames(sum.biomass.demographic) <- c("Juvenile Female", "Juvenile Male", "Adult Female", "Adult Male")
kable(sum.biomass.demographic, caption= "Summer biomass consumed by demographic (in kgs)")


# Total biomass for the entire population in Summer (metric tons)
Summer_tot_biomass <- (tot_biomass_SummerJuvF + tot_biomass_SummerJuvM + tot_biomass_SummerAdF + tot_biomass_SummerAdM) / 1000
Summer_tot_biomass_sum <- get_biomass_quantiles(Summer_tot_biomass)

kable(Summer_tot_biomass_sum, caption = "Total metric tons of prey consumed in summer")
```
:::

## **Results: Total biomass (metric tons) consumed**

```{r overall biomass calc, cache=TRUE}
total_biomass <- (Summer_tot_biomass + Spring_tot_biomass + Winter_tot_biomass)
overall_biomass_consumed <- as.data.frame(quantile(total_biomass,probs = c(0.5, 0.025, 0.975, 0.25,0.75)))

colnames(overall_biomass_consumed) <- c("Prey consumed (metric tons)")

overall_biomass_consumed <- apply(overall_biomass_consumed, 2, function(x) formatC(x, format = "f", big.mark = ",", digits = 1))

kable(overall_biomass_consumed, caption = "Metric tons of prey consumed December-August")

total_biomass_seasonal_avg_dat <- cbind(Winter_tot_biomass_sum, Spring_tot_biomass_sum, Summer_tot_biomass_sum) %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  pivot_longer(cols = -rowname, names_to = "season", values_to = "value") %>%
  pivot_wider(names_from = rowname, values_from = value)

fig4dat <- total_biomass_seasonal_avg_dat %>%
  mutate(
    avg = `50%`,
    ci_min = `2.5%`,
    ci_max = `97.5%`,
    ci_low = `25%`,
    ci_high = `75%`,
    season = factor(season, levels = c('Winter_tot_biomass_sum', 'Spring_tot_biomass_sum', 'Summer_tot_biomass_sum'),
                    labels = c('Winter', 'Spring', 'Summer')))

total_biomass_seasonal_avg_dat_formatted <- total_biomass_seasonal_avg_dat %>%
  mutate_all(~ formatC(., format = "f", big.mark = ",", digits = 1))

kable(total_biomass_seasonal_avg_dat_formatted, caption = "Biomass (in metric tons) consumed by season")
```

### Figure 4: Total Biomass Consumed
```{r reformatting for figure, cache=TRUE}
# Rename columns and factor the seasons
seasonorder <- c('Winter', 'Spring', 'Summer')

fig4plot <- ggplot(data=fig4dat, aes(x=factor(season, level= seasonorder), y=avg))+
  geom_bar(stat="identity", fill = "grey", size = 0.5,color = "black")+
  geom_errorbar(data=fig4dat, aes(ymin = ci_min, ymax = ci_max), width =0.1)+
  geom_errorbar(data=fig4dat, aes(ymin = ci_low, ymax = ci_high), width =0.0, 
                linewidth = 6, col = "#E2A43B")+ 
  theme_light()  + 
  theme(text=element_text(size=12)) + 
  xlab('Season') + 
  ylab('Total Biomass of Prey Consumed (metric tons)')

fig4plot

ggsave("Figures/Figure4.pdf", fig4plot, device = "pdf", units = c("in"), width = 6, height = 4, dpi = 300)

```


# ZOIB model for Chinook Diet Proportions

## Data setup 
```{r data set up for ZOIB, cache=TRUE}
set.seed(99)

RRAdat <- read_xlsx("model inputs.xlsx", sheet = "Chinook RRA")
RRAdat$PRCNT <- RRAdat$RRA

chinook <- RRAdat %>% select(sample, PRCNT) 

metadata <- read_xlsx("model inputs.xlsx", sheet = "metadata")

model_df <- merge(chinook, metadata, all = TRUE, by.x = "sample") %>% replace_na(list(PRCNT=0))

model_df$species <- "ONCTSH"

## Define useful variables
n_samples<- length(unique(model_df$sample)) # Number of samples in the datset
sample_list<- unique(model_df$sample) # List of sample IDs in dataset
target_spp <- "ONCTSH"

## Build the structure of the model df
model_df$sample<- as.factor(model_df$sample)
model_df$site<- as.factor(model_df$site)
model_df$season<- as.factor(model_df$season)
```

## Plot of sample diet proportions, by season and sex

```{r chinook raw data plots, cache=TRUE}
plot_col<- viridis(5)# Colors
site_cols<- c("Tatoosh Cut"=plot_col[1], "Sea Lion Rock"=plot_col[3])
site_cols_fill<- c("Tatoosh Cut"=alpha(plot_col[1], 0.25), "Sea Lion Rock"=alpha(plot_col[3], 0.25))

# Labels
site_labs<- c("Tatoosh Cut"="Tatoosh Cut", "Sea Lion Rock"="Sea Lion Rock") 
season_labs<- c("winter"="Winter", "spring"="Spring", "summer"="Summer")
season_labels_alt<- c("W", "Spr", "Sum")

# Build plot
prop_plot1<- 
  ggplot(model_df, aes(x=season, y=PRCNT, fill = site, color = site)) + 
  geom_jitter(alpha=0.25, shape=16, width = 0.25) +  
  geom_violin(aes(), trim = TRUE, scale = "area") +
  labs(x="Month", y="Diet proportion") +
   scale_x_discrete(labels = season_labs) +  
  scale_fill_manual(values=site_cols_fill) + 
  scale_color_manual(values=site_cols) + 
  theme_bw() +
  ggtitle("Sample derived Chinook diet proportions") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_text(size=12, color="black",                               face="bold"), strip.text.y = element_text(size=12, color="black", face="bold"), legend.position = c(0.8, 0.8), 
        axis.title.x = element_text(face="bold", size=14, hjust = 0.5, vjust = 0.00),
        axis.title.y = element_text(face="bold", size=14, hjust = 0.5, vjust = 2.50),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        panel.spacing = unit(1.15, "lines")) + guides(fill = guide_legend(title = NULL), color = guide_legend(title = NULL))


prop_plot2 <- ggplot(model_df, aes(x = season, y = PRCNT)) + 
  geom_jitter(alpha = 0.25, shape = 16, width = 0.25) +  
  geom_violin(trim = TRUE, scale = "area", fill = "grey", color = "black") +  # Combine sites
  labs(x = "Month", y = "Diet proportion") +
  scale_x_discrete(labels = season_labs) +  
  theme_bw() +
  ggtitle("Sample derived Chinook diet proportions") + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    strip.text.x = element_text(size = 12, color = "black", face = "bold"), 
    strip.text.y = element_text(size = 12, color = "black", face = "bold"), 
    legend.position = c(0.8, 0.8), 
    axis.title.x = element_text(face = "bold", size = 14, hjust = 0.5, vjust = 0.00),
    axis.title.y = element_text(face = "bold", size = 14, hjust = 0.5, vjust = 2.50),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    panel.spacing = unit(1.15, "lines")
  )

prop_plot1
prop_plot2

```

## Fit the model

Here we run 4 MCMC chains on a seasonal model that assumes that the average diet proportion varies by season.

```{r ZOIB model fit, cache=TRUE}
chains = 4 # Number of MCMC chains to run

seasonal<- zoib(model_df$PRCNT ~ as.factor(season)|1|1, data=model_df, random = 0, EUID=model_df$sample, joint=FALSE, 
                zero.inflation = TRUE, one.inflation = FALSE, 
                n.iter = 2000, 
                n.chain = chains, 
                n.thin = 1, 
                n.burn = 1000)

preds_seasonal<- rbind(seasonal$ypred[[1]], seasonal$ypred[[2]], seasonal$ypred[[3]], seasonal$ypred[[4]])
```

## Produce n samples from the posterior predictive distributions, for each season

```{r ZOIB sample replicates, cache=TRUE}
n_samples<- 2000
# This is the start of each season in the sample list
winter_start <- as.numeric(row.names(model_df[model_df$season=="winter", ]))[1]
spring_start<- as.numeric(row.names(model_df[model_df$season=="spring", ]))[1]
summer_start<- as.numeric(row.names(model_df[model_df$season=="summer", ]))[1]

# Generate n samples for each seasonal estimate
winter_samples <- sample(preds_seasonal[,winter_start], n_samples, TRUE)
summary(winter_samples)
spring_samples<- sample(preds_seasonal[,spring_start], n_samples, TRUE)
summer_samples<- sample(preds_seasonal[,summer_start], n_samples, TRUE)
```

## Plot seasonal diet proportions (histogram of n_samples)

```{r Plot ZOIB samples, cache=TRUE}
par(mfrow=c(3,1), oma=c(5,4,2,1), mar=c(1,0,1,0)+0.1)
hist(winter_samples, xlim=c(0,0.3), ann=FALSE, yaxt="n", col = alpha(plot_col[1], 0.50), breaks=15)
mtext(side=2, font=2, line=2.0, "Winter", outer = FALSE)
hist(spring_samples, xlim=c(0,0.3), ann=FALSE, yaxt="n", col = alpha(plot_col[2], 0.50), breaks=15)
mtext(side=2, font=2, line=2.0, "Spring", outer = FALSE)
hist(summer_samples, xlim=c(0,0.3), ann=FALSE, yaxt="n", col = alpha(plot_col[3], 0.50), breaks=15)
mtext(side=2, font=2, line=2.0, "Summer", outer = FALSE)
mtext(side=1, font=2, line=2.0, "Diet proportion", outer = TRUE)
```

# Total Chinook biomass consumed

## Chinook diet percentage from ZOIB model

::: panel-tabset
### Overall

```{r overall Chinook diet proportion, cache=TRUE}
chinprop <- cbind(winter_samples, spring_samples, summer_samples)

chinprop_CI <- quantile(chinprop,probs = c(0.5, 0.025, 0.975, 0.25,0.75))

kable(chinprop_CI)

```

### Winter

```{r winter Chinook diet proportion, cache=TRUE}
winter_allChinook_proportion <- quantile(winter_samples,probs = c(0.5, 0.025, 0.975, 0.25,0.75))
winter_allChinook_proportion
```

### Spring

```{r spring Chinook diet proportion, cache=TRUE}
spring_allChinook_proportion <- quantile(spring_samples,probs = c(0.5, 0.025, 0.975, 0.25,0.75))
spring_allChinook_proportion
```

### Summer

```{r summer Chinook diet proportion, cache=TRUE}
summer_allChinook_proportion <- quantile(summer_samples,probs = c(0.5, 0.025, 0.975, 0.25,0.75))
summer_allChinook_proportion
```
:::

## Seasonal consumption of Chinook Biomass  (in metric tons) 

::: panel-tabset
### Winter

```{r winter Chinook biomass consumed, cache=TRUE}
chin_mass_calc <- function(tot_biomass, sample_list) {
  total_chin <- tot_biomass * sample_list
  quantiles <- quantile(total_chin, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
  result <- list(
    total_samples = total_chin,
    quantiles = quantiles)
   return(result)
 }

#by demographic
Win.Juv.F_tot_chin <- chin_mass_calc(tot_biomass_WinterJuvF, winter_samples)
Win.Juv.M_tot_chin <- chin_mass_calc(tot_biomass_WinterJuvM, winter_samples)
Win.Ad.F_tot_chin <- chin_mass_calc(tot_biomass_WinterAdF, winter_samples)
Win.Ad.M_tot_chin <- chin_mass_calc(tot_biomass_WinterAdM, winter_samples)

# Calculate total for Winter 
Winter_tot_chin <- (Win.Juv.F_tot_chin$total_samples + 
                    Win.Juv.M_tot_chin$total_samples + 
                    Win.Ad.F_tot_chin$total_samples + 
                    Win.Ad.M_tot_chin$total_samples) / 1000

Winter_tot_chin_sum <- quantile(Winter_tot_chin, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
Winter_tot_chin_sum

```

### Spring

```{r spring Chinook biomass consumed, cache=TRUE}
#by demographic
Spr.Juv.F_tot_chin <- chin_mass_calc(tot_biomass_SpringJuvF, spring_samples)
Spr.Juv.M_tot_chin <- chin_mass_calc(tot_biomass_SpringJuvM, spring_samples)
Spr.Ad.F_tot_chin <- chin_mass_calc(tot_biomass_SpringAdF, spring_samples)
Spr.Ad.M_tot_chin <- chin_mass_calc(tot_biomass_SpringAdM, spring_samples)

# Calculate total for spring
Spring_tot_chin <- (Spr.Juv.F_tot_chin$total_samples + 
                    Spr.Juv.M_tot_chin$total_samples + 
                    Spr.Ad.F_tot_chin$total_samples + 
                    Spr.Ad.M_tot_chin$total_samples) / 1000

Spring_tot_chin_sum <- quantile(Spring_tot_chin, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
Spring_tot_chin_sum
```

### Summer

```{r summer Chinook biomass consumed, cache=TRUE}
#by demographic
Sum.Juv.F_tot_chin <- chin_mass_calc(tot_biomass_SummerJuvF, summer_samples)
Sum.Juv.M_tot_chin <- chin_mass_calc(tot_biomass_SummerJuvM, summer_samples)
Sum.Ad.F_tot_chin <- chin_mass_calc(tot_biomass_SummerAdF, summer_samples)
Sum.Ad.M_tot_chin <- chin_mass_calc(tot_biomass_SummerAdM, summer_samples)

# Calculate total for summer
Summer_tot_chin <- (Sum.Juv.F_tot_chin$total_samples + 
                    Sum.Juv.M_tot_chin$total_samples + 
                    Sum.Ad.F_tot_chin$total_samples + 
                    Sum.Ad.M_tot_chin$total_samples) / 1000

Summer_tot_chin_sum <- quantile(Summer_tot_chin, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
Summer_tot_chin_sum
```
:::

## Summary of Chinook Biomass (in metric tons) consumed by SSL

```{r total Chinook biomass, cache=TRUE}
Total_biomass_Chinook <- (Winter_tot_chin + Spring_tot_chin + Summer_tot_chin)
Total_biomass_Chinook_summary <- quantile(Total_biomass_Chinook,probs = c(0.5, 0.025, 0.975, 0.25,0.75))
Total_biomass_Chinook_summary
```

```{r summary of Chinook biomass consumed, cache=TRUE}
tot_chin_data <- list(
  Winter = Winter_tot_chin_sum,
  Spring = Spring_tot_chin_sum,
  Summer = Summer_tot_chin_sum
)

# Convert the list to a data frame
biomass_chinook_seasonal_df <- bind_rows(
  lapply(names(tot_chin_data), function(season) {
    data.frame(
      season = season,
      avg = tot_chin_data[[season]][["50%"]],
      ci_min = tot_chin_data[[season]][["2.5%"]],
      ci_max = tot_chin_data[[season]][["97.5%"]],
      ci_low = tot_chin_data[[season]][["25%"]],
      ci_high = tot_chin_data[[season]][["75%"]]
    )
  })
)

biomass_chinook_seasonal_df

write.csv(biomass_chinook_seasonal_df, file="code outputs/biomass_allChinook.csv")
```



# Part 4: Consumption of Age-0 Chinook

## Adding uncertainty around age-0 proportions

Here we use a beta distribution to add uncertainty based off of the proportion of age-0 fish per season, and the total number of samples collected each season. Further, we calculate the values used later in the sensitivity analysis.

### Age-0 proportion of Chinook recovered

```{r age0 prop calculation, cache=TRUE}
proportions <- c(winter = 0.58, spring = 0.55, summer = 0.35)
sample_sizes <- c(winter = 29, spring = 18, summer = 17)

# Define percentage changes for sensitivity analysis
percent_changes <- c(0.10, 0.25, 0.50)

# Function to adjust proportions and include original proportions
adjust_proportions <- function(proportions, percent_changes) {
  # Create a data frame to store the original proportions
  df <- data.frame(Season = names(proportions), 
                   Original_Proportion = proportions)
  
  # Loop over each percent change to calculate adjusted proportions
  for (change in percent_changes) {
    # Calculate adjusted proportions (Up and Down) and add them to the data frame
    df[[paste0("Change_", change * 100, "%_Up")]] <- sapply(proportions, function(prop) prop * (1 + change))
    df[[paste0("Change_", change * 100, "%_Down")]] <- sapply(proportions, function(prop) prop * (1 - change))
  }
  
  # Return the data frame with original and adjusted proportions
  df
}

# Apply the function and store the result
proportions_df <- adjust_proportions(proportions, percent_changes)

# Function to generate samples
generate_samples <- function(proportion, sample_size, num_samples) {
  alpha <- proportion * (sample_size - 1) + 1
  beta <- (1 - proportion) * (sample_size - 1) + 1
  rbeta(num_samples, alpha, beta)
}

# Number of samples to generate
num_samples <- 2000

# Function to calculate summary statistics
calculate_summary_stats <- function(samples) {
  summary_x <- summary(samples)
  mean_x <- mean(samples)
  cv <- sd(samples) / mean_x
  data.frame(Mean = mean_x, SD = sd(samples), CV = cv, Min = summary_x[["Min."]], 
             `1st Qu.` = summary_x[["1st Qu."]], Median = summary_x[["Median"]], 
             Mean_Summary = summary_x[["Mean"]], `3rd Qu.` = summary_x[["3rd Qu."]], 
             Max = summary_x[["Max."]])
}

# Initialize list to store all results and samples
all_results <- list()
all_samples <- list()

# Iterate over each row in the proportions data frame
for (i in 1:nrow(proportions_df)) {
  # Extract the season and its sample size
  season <- proportions_df$Season[i]
  sample_size <- sample_sizes[season]
  
  # Iterate over each adjusted proportion (original and adjusted)
  for (j in 2:ncol(proportions_df)) {
    change_name <- colnames(proportions_df)[j]
    proportion <- proportions_df[i, j]
    
    # Generate samples for each adjusted proportion
    samples <- generate_samples(proportion, sample_size, num_samples)
    
    # Calculate summary statistics
    summary_stats <- calculate_summary_stats(samples)
    
    # Store results and samples
    all_results[[paste(season, change_name, sep = "_")]] <- summary_stats
    all_samples[[paste(season, change_name, sep = "_")]] <- samples
    
    # Convert to a data frame for visualization
    samples_df <- data.frame(
      Proportion = samples,
      Season = season,
      Change = change_name
    )
    
    # Plot the distribution of generated proportions for each season and change
  }
}
# Collect results into a single data frame for comparison
results_df <- do.call(rbind, lapply(names(all_results), function(change_name) {
  stats <- all_results[[change_name]]
  stats_df <- as.data.frame(stats)  # Convert to a data frame
  stats_df$Change <- change_name
  stats_df
}))

# Print the cleaned-up results
kable(results_df)

# Function to retrieve generated samples for a specific season and change
get_samples <- function(season, change_name) {
  sample_key <- paste(season, change_name, sep = "_")
  all_samples[[sample_key]]
}
```


```{r plot ratio of age-0 by season, cache=TRUE}
# Initialize a data frame to store all samples for plotting
# Initialize a data frame to store all samples for plotting
combined_samples_df <- data.frame()

# Initialize a data frame to store means for plotting
means_df <- data.frame()

# Iterate over each row in the proportions data frame
for (i in 1:nrow(proportions_df)) {
  # Extract the season and its sample size
  season <- proportions_df$Season[i]
  sample_size <- sample_sizes[season]
  
  # Iterate over each adjusted proportion (original and adjusted)
  for (j in 2:ncol(proportions_df)) {
    change_name <- colnames(proportions_df)[j]
    proportion <- proportions_df[i, j]
    
    # Generate samples for each adjusted proportion
    samples <- generate_samples(proportion, sample_size, num_samples)
    
    # Calculate mean for each group and store it in the means_df
    mean_value <- mean(samples)
    means_df <- rbind(means_df, data.frame(Proportion = mean_value, Season = season, Change = change_name))
    
    # Convert to a data frame for visualization and store in the combined data frame
    samples_df <- data.frame(
      Proportion = samples,
      Season = season,
      Change = change_name
    )
    combined_samples_df <- rbind(combined_samples_df, samples_df)
  }
}

# Plot the combined samples for each season with adjustments overlaid
plot_combined_overlay <- ggplot(combined_samples_df, aes(x = Proportion, fill = Change)) +
  geom_histogram(position = "identity", binwidth = 0.02, alpha = 0.5) +  # Use 'identity' to overlay histograms
  geom_vline(data = means_df, aes(xintercept = Proportion, color = Change), linetype = "dashed", size = 1) +  # Add mean lines
  labs(title = "Ratio of Age-0 to Other Aged Fish with Adjustments",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("Original_Proportion" = "gray", 
                               "Change_10%_Up" = "#1f77b4", 
                               "Change_10%_Down" = "#2ca02c", 
                               "Change_25%_Up" = "#ff7f0e", 
                               "Change_25%_Down" = "#d62728", 
                               "Change_50%_Up" = "#9467bd", 
                               "Change_50%_Down" = "#8c564b")) +
  scale_color_manual(values = c("Original_Proportion" = "black", 
                                "Change_10%_Up" = "#1f77b4", 
                                "Change_10%_Down" = "#2ca02c", 
                                "Change_25%_Up" = "#ff7f0e", 
                                "Change_25%_Down" = "#d62728", 
                                "Change_50%_Up" = "#9467bd", 
                                "Change_50%_Down" = "#8c564b")) +  # Matching colors for mean lines
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank()) +
  facet_wrap(~ Season, ncol = 1)  # Facet by season to create three plots (one for each season)


plot_combined_overlay
```




```{r by percent change, cache=TRUE}
# Plot the combined samples with faceting by 'Change'
# Filter out 'Original_Proportion' from the combined samples data frame for the faceted plot
faceted_samples_df <- combined_samples_df[combined_samples_df$Change != "Original_Proportion", ]

# Create the faceted plot without the "Original_Proportion"
plot_combined <- ggplot(faceted_samples_df, aes(x = Proportion, fill = Season)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Ratio of Age-0 to Other Aged Fish for Different Changes",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("winter" = "#1f77b4", "spring" = "#2ca02c", "summer" = "#ff7f0e")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank()) +
  facet_wrap(~ Change, ncol = 2)

original_samples_df <- combined_samples_df[combined_samples_df$Change == "Original_Proportion", ]

plot_original <- ggplot(original_samples_df, aes(x = Proportion, fill = Season)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Ratio of Age-0 to Other Aged Fish - Original Proportion",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("winter" = "#1f77b4", "spring" = "#2ca02c", "summer" = "#ff7f0e")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank())

plot_combined
plot_original

```

## Determining Age-0 diet proportion of total diet

```{r age0 diet comp calculation , cache=TRUE}
# Original season samples list
season_samples <- list(
  winter = winter_samples,
  spring = spring_samples,
  summer = summer_samples
)

# Initialize list to store all adjusted samples after multiplying by seasonal samples
all_adjusted_samples <- list()

# Iterate over each combination of season and proportion change in the all_samples list
for (sample_key in names(all_samples)) {
  # Extract the season and change details from the sample key (e.g., "winter_Change_10%_Up")
  key_split <- strsplit(sample_key, "_")[[1]]
  season <- key_split[1]
  
  # This captures both the percentage change and whether it's "Up" or "Down" (e.g., "Change_10%_Up")
  change_name <- paste(key_split[2], key_split[3], key_split[4], sep = "_")  
  
  # Multiply the corresponding samples by the seasonal samples for that season
  adjusted_samples <- all_samples[[sample_key]] * season_samples[[season]]
  
  # Store the adjusted samples back into the new list
  all_adjusted_samples[[sample_key]] <- adjusted_samples
}

# Create a combined list to store all samples for each change group across all seasons
overall_samples <- list()

# Combine samples across all seasons for each change group
for (change_group in unique(sapply(names(all_samples), function(sample_key) {
  key_split <- strsplit(sample_key, "_")[[1]]
  paste(key_split[2], key_split[3], key_split[4], sep = "_")
}))) {
  # Collect samples from all seasons for this change group
  combined_samples <- unlist(lapply(names(all_adjusted_samples), function(sample_key) {
    if (grepl(change_group, sample_key)) {
      return(all_adjusted_samples[[sample_key]])
    } else {
      return(NULL)
    }
  }))
  
  # Store the combined samples
  overall_samples[[change_group]] <- combined_samples
}

# Function to retrieve adjusted samples for a specific season and change
get_adjusted_samples <- function(season, change_name) {
  sample_key <- paste(season, change_name, sep = "_")
  all_adjusted_samples[[sample_key]]
}

# Create a summary dataframe for adjusted samples
# Create a summary dataframe for adjusted samples
adjusted_samples_df <- do.call(rbind, lapply(names(all_adjusted_samples), function(sample_key) {
  key_split <- strsplit(sample_key, "_")[[1]]
  season <- key_split[1]
  
  # Correctly capture the full "Change" (e.g., "Change_10%_Up")
  change <- paste(key_split[2], key_split[3], key_split[4], sep = "_")
  
  # Calculate summary statistics including CI
  median_val <- median(all_adjusted_samples[[sample_key]])
  sd_val <- sd(all_adjusted_samples[[sample_key]])
  ci_95 <- quantile(all_adjusted_samples[[sample_key]], probs = c(0.025, 0.975), names = FALSE)
  ci_50 <- quantile(all_adjusted_samples[[sample_key]], probs = c(0.25, 0.75), names = FALSE)
  
  data.frame(
    Season = season,
    Change = change,  # Now includes "Up" or "Down"
    Median = median_val,
    SD = sd_val,
    CI_95_Lower = ci_95[1],
    CI_95_Upper = ci_95[2],
    CI_50_Lower = ci_50[1],
    CI_50_Upper = ci_50[2]
  )
}))

# Create a summary dataframe for overall samples
overall_samples_df <- do.call(rbind, lapply(names(overall_samples), function(change_group) {
  # Calculate summary statistics for the combined overall samples
  median_val <- median(overall_samples[[change_group]])
  sd_val <- sd(overall_samples[[change_group]])
  ci_95 <- quantile(overall_samples[[change_group]], probs = c(0.025, 0.975), names = FALSE)
  ci_50 <- quantile(overall_samples[[change_group]], probs = c(0.25, 0.75), names = FALSE)
  
  data.frame(
    Season = "Overall",
    Change = change_group,
    Median = median_val,
    SD = sd_val,
    CI_95_Lower = ci_95[1],
    CI_95_Upper = ci_95[2],
    CI_50_Lower = ci_50[1],
    CI_50_Upper = ci_50[2]
  )
}))

# Combine the seasonal and overall dataframes
final_summary_df <- rbind(adjusted_samples_df, overall_samples_df)

# Display the final summary dataframe
kable(final_summary_df)


```

```{r plot age-0 sensitivity by season, cache=TRUE}
combined_samples_df <- do.call(rbind, lapply(names(all_adjusted_samples), function(sample_key) {
  key_split <- strsplit(sample_key, "_")[[1]]
  season <- key_split[1]
  change <- paste(key_split[2], key_split[3], key_split[4], sep = "_")
  
  # Create a dataframe for each set of adjusted samples
  data.frame(
    Proportion = all_adjusted_samples[[sample_key]],
    Season = season,
    Change = change
  )
}))

# Use the adjusted_samples_df for the mean lines
means_df <- adjusted_samples_df

# Create the plot with adjusted data for each season
plot_combined_overlay <- ggplot(combined_samples_df, aes(x = Proportion, fill = Change)) +
  geom_histogram(position = "identity", binwidth = 0.001, alpha = 0.5) +  # Overlay histograms
  labs(title = "Proportion Adjustments by Season",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("Change_10%_Down" = "#66c2a5", 
                               "Change_10%_Up" = "#3288bd", 
                               "Change_25%_Down" = "#abdda4", 
                               "Change_25%_Up" = "#fdae61", 
                               "Change_50%_Down" = "#e6f598", 
                               "Change_50%_Up" = "#d53e4f", 
                               "Original_Proportion" = "#542788")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank()) +
  facet_wrap(~ Season, ncol = 1)  # Facet by season

# Add mean lines for each group using `means_df`
plot_combined_overlay <- plot_combined_overlay + 
  geom_vline(data = means_df, aes(xintercept = Mean, color = Change), 
             linetype = "dashed", size = 1) +
  scale_color_manual(values = c("Change_10%_Down" = "#66c2a5", 
                                "Change_10%_Up" = "#3288bd", 
                                "Change_25%_Down" = "#abdda4", 
                                "Change_25%_Up" = "#fdae61", 
                                "Change_50%_Down" = "#e6f598", 
                                "Change_50%_Up" = "#d53e4f", 
                                "Original_Proportion" = "#542788"))  # Match colors to fill
plot_combined_overlay
```

```{r plot of age0 diet comp, cache=TRUE}
# Filter out the "Original_Proportion_NA" if it exists in the adjusted_samples_df
faceted_samples_df <- combined_samples_df[combined_samples_df$Change != "Original_Proportion_NA", ]

plot_combined1 <- ggplot(faceted_samples_df, aes(x = Proportion, fill = Season)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Proportion of Chinook in diet",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("winter" = "#1f77b4", "spring" = "#2ca02c", "summer" = "#ff7f0e")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank()) +
  facet_wrap(~ Change, ncol = 2)

# Plot for "Original_Proportion"
original_samples_df <- combined_samples_df[combined_samples_df$Change == "Original_Proportion_NA", ]

plot_original2 <- ggplot(original_samples_df, aes(x = Proportion, fill = Season)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Ratio of Age-0 to Other Aged Fish - Original Proportion",
       x = "Proportion",
       y = "Frequency") +
  scale_fill_manual(values = c("winter" = "#1f77b4", "spring" = "#2ca02c", "summer" = "#ff7f0e")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right", legend.title = element_blank())

plot_combined1
plot_original2

```

```{r, cache=TRUE}

age0_model_df <- original_samples_df
age0_model_df$group <- "Age-0"

chinprop <- as.data.frame(chinprop)

chinMODELdf <- pivot_longer(chinprop, cols = 1:3, names_to = "season", values_to = "PRCNT")
chinMODELdf$group <- "All"


age0_model_df$season <- factor(age0_model_df$Season, levels =  c('winter', 'spring', 'summer'), labels = c('Winter', 'Spring', 'Summer'))

chinMODELdf$season <- factor(chinMODELdf$season, levels =  c('winter_samples', 'spring_samples', 'summer_samples'), labels = c('Winter', 'Spring', 'Summer'))

# Combine both data frames into one
proportion_df_plot <- bind_rows(
  age0_model_df %>%
    mutate(value = Proportion, group = "Age-0"),
  chinMODELdf %>%
    rename(value = PRCNT)
)

proportion_df_plot$group <- factor(proportion_df_plot$group, levels = c("All", "Age-0"))
# Use a single ggplot object with facet_wrap
combined_plot <- ggplot(proportion_df_plot, aes(x = season, y = value)) +
  geom_jitter(aes(color = 2, alpha = 0.1)) +
  geom_boxplot(fill = 'lightgrey') +
  labs(x = "Data Source", y = "Chinook Diet proportion (RRA)") + 
  facet_wrap(~ group, ncol = 1, strip.position = "right") +  # Position labels on the right
  scale_fill_manual(values = grey) + 
  theme_bw() +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    strip.text = element_text(size = 12, color = "black", face = "bold"), 
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold", size = 14, hjust = 0.5, vjust = 2.5),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    panel.spacing = unit(1.15, "lines"),
    strip.placement = "outside"  # Place facet labels outside the plot area
  )


combined_plot

ggsave("Figures/Figure 5.pdf", combined_plot, units = "in", width = 8, height = 5, dpi = 300)
```

### Extract original proportion samples

```{r age-0 diet proportion, cache=TRUE}
winter_age0_og <- get_adjusted_samples("winter", "Original_Proportion")
quantile(winter_age0_og,probs = c(0.5, 0.025, 0.975, 0.25,0.75))

spring_age0_og <- get_adjusted_samples("spring", "Original_Proportion")
quantile(spring_age0_og,probs = c(0.5, 0.025, 0.975, 0.25,0.75))
summer_age0_og <- get_adjusted_samples("summer", "Original_Proportion")
quantile(summer_age0_og,probs = c(0.5, 0.025, 0.975, 0.25,0.75))

chinprop_age0 <- cbind(winter_age0_og, spring_age0_og, summer_age0_og)

chinprop_CI_age0 <- quantile(chinprop_age0,probs = c(0.5, 0.025, 0.975, 0.25,0.75))

kable(chinprop_CI_age0)


```

# Function to calculate age-0 biomass (metric tons) for each season and adjusted proportion

```{r age-0 biomass function, cache=TRUE}
# Initialize an empty list to store the adjusted samples
adjusted_samples_list <- list()

# Initialize an empty data frame to store the seasonal biomass results
seasonal_biomass_samples_df <- data.frame()

# Initialize an empty data frame to store the total biomass results across all seasons
total_biomass_samples_df <- data.frame()

# Function to calculate biomass for a given season
calculate_biomass <- function(season, current_samples) {
  results <- list()
  
  if (season == "winter") {
    results$Win.Juv.F_age0bm <- chin_mass_calc(tot_biomass_WinterJuvF, current_samples)
    results$Win.Juv.M_age0bm <- chin_mass_calc(tot_biomass_WinterJuvM, current_samples)
    results$Win.Ad.F_age0bm <- chin_mass_calc(tot_biomass_WinterAdF, current_samples)
    results$Win.Ad.M_age0bm <- chin_mass_calc(tot_biomass_WinterAdM, current_samples)
    
    # Calculate total biomass for Winter
    results$total_biomass <- (results$Win.Juv.F_age0bm$total_samples +
                              results$Win.Juv.M_age0bm$total_samples +
                              results$Win.Ad.F_age0bm$total_samples +
                              results$Win.Ad.M_age0bm$total_samples) / 1000
  } else if (season == "spring") {
    results$Spr.Juv.F_age0bm <- chin_mass_calc(tot_biomass_SpringJuvF, current_samples)
    results$Spr.Juv.M_age0bm <- chin_mass_calc(tot_biomass_SpringJuvM, current_samples)
    results$Spr.Ad.F_age0bm <- chin_mass_calc(tot_biomass_SpringAdF, current_samples)
    results$Spr.Ad.M_age0bm <- chin_mass_calc(tot_biomass_SpringAdM, current_samples)
    
    # Calculate total biomass for Spring
    results$total_biomass <- (results$Spr.Juv.F_age0bm$total_samples +
                              results$Spr.Juv.M_age0bm$total_samples +
                              results$Spr.Ad.F_age0bm$total_samples +
                              results$Spr.Ad.M_age0bm$total_samples) / 1000
  } else if (season == "summer") {
    results$Sum.Juv.F_age0bm <- chin_mass_calc(tot_biomass_SummerJuvF, current_samples)
    results$Sum.Juv.M_age0bm <- chin_mass_calc(tot_biomass_SummerJuvM, current_samples)
    results$Sum.Ad.F_age0bm <- chin_mass_calc(tot_biomass_SummerAdF, current_samples)
    results$Sum.Ad.M_age0bm <- chin_mass_calc(tot_biomass_SummerAdM, current_samples)
    
    # Calculate total biomass for Summer
    results$total_biomass <- (results$Sum.Juv.F_age0bm$total_samples +
                              results$Sum.Juv.M_age0bm$total_samples +
                              results$Sum.Ad.F_age0bm$total_samples +
                              results$Sum.Ad.M_age0bm$total_samples) / 1000
  }
  
  return(results)
}

# Initialize lists to store results
all_demographic_results <- list()
seasonal_biomass_samples_df <- data.frame()
total_biomass_samples_df <- data.frame()
total_biomass_across_seasons <- list()

# Loop over all generated samples for each season and change
for (sample_key in names(all_samples)) {
  key_parts <- strsplit(sample_key, "_")[[1]]
  season <- key_parts[1]
  change_name <- paste(key_parts[-1], collapse = "_")

  current_samples <- all_samples[[sample_key]]
  
  if (season == "winter") {
    adjusted_samples <- winter_samples * current_samples
  } else if (season == "spring") {
    adjusted_samples <- spring_samples * current_samples
  } else if (season == "summer") {
    adjusted_samples <- summer_samples * current_samples
  }
  
  adjusted_samples_list[[sample_key]] <- adjusted_samples
  
  # Perform the biomass calculation
  results <- calculate_biomass(season, adjusted_samples)
  
  # Store all demographic results
  all_demographic_results[[sample_key]] <- results
  
  # Store the seasonal biomass
  seasonal_biomass_samples_df <- rbind(seasonal_biomass_samples_df, 
                                       data.frame(Season = season, 
                                                  Change = change_name, 
                                                  Seasonal_Samples = results$total_biomass))

  # Accumulate total biomass across seasons
  if (!is.null(total_biomass_across_seasons[[change_name]])) {
    total_biomass_across_seasons[[change_name]] <- total_biomass_across_seasons[[change_name]] + results$total_biomass
  } else {
    total_biomass_across_seasons[[change_name]] <- results$total_biomass
  }
}

# Convert total biomass across seasons list into a data frame
for (change_name in names(total_biomass_across_seasons)) {
  total_biomass_samples_df <- rbind(total_biomass_samples_df, 
                                    data.frame(Change = change_name, 
                                               Total_Samples = total_biomass_across_seasons[[change_name]]))
}


# Summarize the seasonal biomass results
seasonal_biomass_summary <- seasonal_biomass_samples_df %>%
  group_by(Season, Change) %>%
  summarise(
    avg = quantile(Seasonal_Samples, probs = 0.5),
    ci_min = quantile(Seasonal_Samples, probs = 0.025),
    ci_low = quantile(Seasonal_Samples, probs = 0.25),
    ci_high = quantile(Seasonal_Samples, probs = 0.75),
    ci_max = quantile(Seasonal_Samples, probs = 0.975),
    .groups = 'drop'
  )

# Summarize the total biomass results across seasons
age0_biomass_summary <- total_biomass_samples_df %>%
  group_by(Change) %>%
  summarise(avg = quantile(Total_Samples, probs = 0.5),
    ci_min = quantile(Total_Samples, probs = 0.025),
    ci_low = quantile(Total_Samples, probs = 0.25),
    ci_high = quantile(Total_Samples, probs = 0.75),
    ci_max = quantile(Total_Samples, probs = 0.975),
    .groups = 'drop')

age0_biomass_summary$Change_Simplified <- sub(".*_(Original_Proportion|Change_.*)", "\\1", age0_biomass_summary$Change)

age0_biomass_summary_seasonal <- seasonal_biomass_summary %>%
  mutate(Change_Simplified = gsub("_", " ", Change),
         Change_Simplified = gsub("Change", "", Change),
         Change_Simplified = gsub("Up", "Increase", Change),
         Change_Simplified = gsub("Down", "Decrease", Change),
         # Convert Season to a factor with specified levels to ensure order
         Season = factor(Season, levels = c("winter", "spring", "summer")))

changes_df <- age0_biomass_summary_seasonal %>%
  filter(!grepl("Original_Proportion", Change_Simplified))
original_proportions_df <- age0_biomass_summary_seasonal %>%
  filter(grepl("Original_Proportion", Change_Simplified)) %>%
  mutate(
    Season = recode(Season, "winter" = "Winter", "spring" = "Spring", "summer" = "Summer"))


kable(age0_biomass_summary, label = "Age-0 Chinook Biomass Consumed")

```

```{r sensitivity of age-0 biomass plot, cache=TRUE}

# Plot the changes (excluding original proportions) with facet_wrap
change_plot <- ggplot(changes_df, aes(x = Season, y = avg)) +
  geom_bar(stat = "identity", fill = "gray") +
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2, color = "black", size = 0.8) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0, color = "orange", size = 6) +
  labs(title = "Total Biomass of Prey Consumed by Season (Excluding Original Proportions)",
       x = "Season",
       y = "Total Biomass of Prey Consumed (metric tons)") +
  theme_minimal() +
  facet_wrap(~ Change_Simplified, scales = "fixed", ncol = 2)

change_plot


# Plot only the original proportions
original_plot <- ggplot(original_proportions_df, aes(x = Season, y = avg)) +
  geom_bar(stat = "identity", fill = "gray") +
  # Add larger confidence interval (ci_min and ci_max)
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2, color = "black", size = 0.8) +
  # Add smaller confidence interval (ci_low and ci_high)
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0, color = "orange", size = 6) +
  labs(title = "Total Biomass of Prey Consumed - Original Proportions",
       x = "Season",
       y = "Total Biomass of Prey Consumed (metric tons)") +
  theme_minimal()

# Display the original proportions plot
original_plot

kable(original_proportions_df)
```

### Figure 6: Total chinook consumed relative to age-0 consumed
```{r figure 6 biomass plot, cache=TRUE}
allChin <- biomass_chinook_seasonal_df %>% select(season, avg, ci_min, ci_max, ci_low, ci_high)


original_proportions_df$season <- original_proportions_df$Season
age0chin <- original_proportions_df %>% select(season, avg, ci_min, ci_max, ci_low, ci_high)


allChin$group <- "all"
age0chin$group <- "juv"

fig6dat <- rbind(allChin, age0chin) 

fig6dat$group <- factor(fig6dat$group, levels =  c('all', 'juv'), 
                        labels = c('Chinook', 'Age-0 Chinook'))

fig6plot <- ggplot(data=fig6dat, aes(x=factor(season, level= seasonorder), y=avg, fill = group)) +
  geom_bar(stat="identity", size = 0.5, position = "dodge", aes(fill = group), colour = "black") +
  scale_fill_manual(values = c("snow4","snow1"), name = "Age-class") + 
  geom_errorbar(data=fig6dat, aes(ymin = ci_min, ymax = ci_max), width =0.1,  position = position_dodge(width = 0.9))+ 
  geom_errorbar(data=fig6dat, aes(ymin = ci_low, ymax = ci_high), width =0.0, size = 6, col = '#55AFBD', position = position_dodge(width = 0.9)) + 
  theme_light()  +  
  theme(text=element_text(size=12)) + 
  xlab('Season') + 
  ylab('Biomass of Chinook Consumed (metric tons)')

fig6plot

ggsave("Figures/Figure6.pdf", fig6plot, units = c("in"), width = 8, 
       height = 5, dpi = 300)

fig4dat_clean <- fig4dat %>% select(c('season','avg','ci_min','ci_max','ci_low', 'ci_high')) %>% mutate(group = "Total Prey")

biomass_plot <- rbind(fig4dat_clean, fig6dat)

biomass_plot <- ggplot(data=biomass_plot, aes(x=factor(season, level= seasonorder), y=avg, fill = group)) +
  geom_bar(stat="identity", size = 0, position = "dodge", aes(fill = group)) +
  scale_fill_manual(values = c("gray30","gray66","lightgray"), name = "Age-class") + 
  geom_errorbar(data=biomass_plot, aes(ymin = ci_min, ymax = ci_max), width =0.1,  position = position_dodge(width = 0.9))+ 
  geom_errorbar(data=biomass_plot, aes(ymin = ci_low, ymax = ci_high), width =0.0, size = 6, col = '#55AFBD', position = position_dodge(width = 0.9)) + 
  theme_light()  +  
  theme(text=element_text(size=12)) + 
  xlab('Season') + 
  ylab('Biomass of Chinook Consumed (metric tons)')
biomass_plot

```

### Bootstrapping 
```{r sensitivity analysis on biomass function, cache=TRUE}
# Function to run bootstrap analysis for a given original distribution and a comparison
# Function to run bootstrap analysis for a given original distribution and a comparison
run_bootstrap_analysis <- function(original_dist, change_dist, n_sims = 2000) {
  # Calculate observed difference
  dist_2<- log(original_dist)
  dist_1<- log(change_dist)
  
  n_samples<- length(dist_1)

  observed_diff <- mean(dist_1) - mean(dist_2)
  
  # All data
  combined <- c(dist_1, dist_2)
  
  # Initialize parameters for bootstrapping
  n <- length(dist_1)  # Size of each resample (same as original datasets)
  diffs <- numeric(n_sims)  # Vector to store resampled differences
  
  # Bootstrap resampling
  for (i in 1:n_sims) {
    shuffled <- sample(combined, length(combined), replace = FALSE)
    dist_1_resample <- shuffled[1:n]
    dist_2_resample <- shuffled[(n + 1):(2 * n)]
    diffs[i] <- mean(dist_1_resample) - mean(dist_2_resample)
  }
  
  # Calculate the p-value
  p_value <- mean(abs(diffs) >= abs(observed_diff))
  
  # Return observed difference and p-value
  return(c(observed_diff = observed_diff, p_value = p_value))
}

# Initialize empty data frame to store results
age0_biomass_bootresults <- data.frame(Change_Comparison = character(), Observed_Difference = numeric(), P_Value = numeric(), stringsAsFactors = FALSE)

# Get the original proportion samples
original_dist <- total_biomass_samples_df %>% filter(Change == "Original_Proportion") %>% pull(Total_Samples)

# List of changes to compare against the original
changes_to_compare <- c("Change_10%_Up", "Change_10%_Down", "Change_25%_Up", "Change_25%_Down", "Change_50%_Up", "Change_50%_Down")

# Loop through each change and run the bootstrap analysis
for (change in changes_to_compare) {
  comparison_dist <- total_biomass_samples_df %>% filter(Change == change) %>% pull(Total_Samples)
  
  # Check if the comparison_dist is empty
  if (length(comparison_dist) == 0) {
    cat("No data for", change, "\n")
  } else {
    # Run the bootstrap analysis
    result <- run_bootstrap_analysis(original_dist, comparison_dist)
    
    # Add the result to the data frame
    age0_biomass_bootresults <- rbind(age0_biomass_bootresults, data.frame(Change_Comparison = change,
                                         Observed_Difference = result['observed_diff'],
                                         P_Value = result['p_value']))
  }
}

# Print the results


age0_biomass_bootresults$inverse_log_values <- exp(age0_biomass_bootresults$Observed_Difference)

age0_biomass_bootresults$percent_change <- (age0_biomass_bootresults$inverse_log_values-1)*100

age0_biomass_bootresults$base_age0avg <- age0_biomass_summary %>%
  filter(Change == "Original_Proportion") %>%
  pull(avg)

age0_biomass_bootresults$expected_value <- age0_biomass_bootresults$base_age0avg * age0_biomass_bootresults$inverse_log_values

age0_biomass_median <- age0_biomass_summary %>% mutate(Median = paste0(round(avg, 1), " (", round(ci_min, 1), "-", round(ci_max, 1), ")")) %>% rename(c("Change_Comparison" = "Change")) %>% select(Change_Comparison, Median)

age0_biomass_bootresults_table <- merge(age0_biomass_median, age0_biomass_bootresults, by = "Change_Comparison", all = TRUE)

kable(age0_biomass_bootresults_table)

```

```{r biomass plot, cache=TRUE}


age0_biomass_summary$Change <- factor(
  age0_biomass_summary$Change,
  levels = c("Original_Proportion", "Change_10%_Down", "Change_10%_Up", "Change_25%_Down", "Change_25%_Up","Change_50%_Down", "Change_50%_Up"),  # Specify the original order
  labels = c("Base Model", "10% Decrease", "10% Increase", "25% Decrease", "25% Increase","50% Decrease", "50% Increase")  # Specify the new labels
)


age0_biomass_plot <-  ggplot(age0_biomass_summary, aes(x = Change, y = `avg`, fill = Change)) +
  geom_bar(stat = "identity", color = "black", fill = "grey") +
  geom_errorbar(aes(ymin = `ci_min`, ymax = `ci_max`), width = 0.2, color = "black", size = 1) +
  geom_errorbar(aes(ymin = `ci_low`, ymax = `ci_high`), width = 0, color = "#55AFBD", size = 6) +
  labs(
       y = "Biomass consumed (metric tons)", 
       x = "Change Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

age0_biomass_plot

ggsave("Figures/Figure8.pdf", plot = age0_biomass_plot, width = 8, height = 5, dpi = 300)

```
## Number of Age-0 Chinook consumed

### Weight of Chinook salmon

```{r weight calculations, cache=TRUE}
# Load required libraries
library(ggplot2)
library(dplyr)
library(patchwork)

# Length functions for each month
DecJuvChinLength <- function() {
  rlnormTrunc(1, log(236.9987), sdlog = 0.5, min = 100, max = 300)
}

JanJuvChinLength <- function() {
  rlnormTrunc(1, log(175.436), sdlog = 0.5, min = 100, max = 300)
}

FebJuvChinLength <- function() {
  rlnormTrunc(1, log(170.2746), sdlog = 0.5, min = 100, max = 300)
}

MarJuvChinLength <- function() {
  rlnormTrunc(1, log(201.9993), sdlog = 0.5, min = 100, max = 300)
}

AprJuvChinLength <- function() {
  rlnormTrunc(1, log(234.8554), sdlog = 0.5, min = 100, max = 300)
}

MayJuvChinLength <- function() {
  rlnormTrunc(1, log(173.6663), sdlog = 0.5, min = 100, max = 300)
}

JunJuvChinLength <- function() {
  rlnormTrunc(1, log(142.0207), sdlog = 0.5, min = 100, max = 300)
}

JulJuvChinLength <- function() {
  rlnormTrunc(1, log(133.1082), sdlog = 0.5, min = 100, max = 300)
}

AugJuvChinLength <- function() {
  rlnormTrunc(1, log(151.4996), sdlog = 0.5, min = 100, max = 300)
}

# Weight functions using lengths
DecJuvChinWt <- function() {
  (9.61 * ((DecJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

JanJuvChinWt <- function() {
  (9.61 * ((JanJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

FebJuvChinWt <- function() {
  (9.61 * ((FebJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

MarJuvChinWt <- function() {
  (9.61 * ((MarJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

AprJuvChinWt <- function() {
  (9.61 * ((AprJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

MayJuvChinWt <- function() {
  (9.61 * ((MayJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

JunJuvChinWt <- function() {
  (9.61 * ((JunJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

JulJuvChinWt <- function() {
  (9.61 * ((JulJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

AugJuvChinWt <- function() {
  (9.61 * ((AugJuvChinLength() / 10)^3.07)) * (1/1000) * (1/1000)
}

# Generate length and weight samples for each month
length_samples_Dec <- replicate(2000, DecJuvChinLength())
weight_samples_Dec <- replicate(2000, DecJuvChinWt())

length_samples_Jan <- replicate(2000, JanJuvChinLength())
weight_samples_Jan <- replicate(2000, JanJuvChinWt())

length_samples_Feb <- replicate(2000, FebJuvChinLength())
weight_samples_Feb <- replicate(2000, FebJuvChinWt())

length_samples_Mar <- replicate(2000, MarJuvChinLength())
weight_samples_Mar <- replicate(2000, MarJuvChinWt())

length_samples_Apr <- replicate(2000, AprJuvChinLength())
weight_samples_Apr <- replicate(2000, AprJuvChinWt())

length_samples_May <- replicate(2000, MayJuvChinLength())
weight_samples_May <- replicate(2000, MayJuvChinWt())

length_samples_Jun <- replicate(2000, JunJuvChinLength())
weight_samples_Jun <- replicate(2000, JunJuvChinWt())

length_samples_Jul <- replicate(2000, JulJuvChinLength())
weight_samples_Jul <- replicate(2000, JulJuvChinWt())

length_samples_Aug <- replicate(2000, AugJuvChinLength())
weight_samples_Aug <- replicate(2000, AugJuvChinWt())

# Combine length samples into a data frame
length_samples_df <- data.frame(
  month = rep(c("December", "January", "February", "March", "April", "May", "June", "July", "August"), each = 2000),
  length_value = c(length_samples_Dec, length_samples_Jan, length_samples_Feb, length_samples_Mar, 
                   length_samples_Apr, length_samples_May, length_samples_Jun, length_samples_Jul, 
                   length_samples_Aug)
)

length_samples_df$month <- factor(
  length_samples_df$month,
  levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August")
)

# Combine weight samples into a data frame
weight_samples_df <- data.frame(
  month = rep(c("December", "January", "February", "March", "April", "May", "June", "July", "August"), each = 2000),
  weight_value = c(weight_samples_Dec, weight_samples_Jan, weight_samples_Feb, weight_samples_Mar, 
                   weight_samples_Apr, weight_samples_May, weight_samples_Jun, weight_samples_Jul, 
                   weight_samples_Aug)
)



weight_samples_df$month <- factor(
  weight_samples_df$month,
  levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August")
)

mean_length_values <- length_samples_df %>%
  group_by(month) %>%
  summarize(mean_length = mean(length_value))

mean_weight_values <- weight_samples_df %>%
  group_by(month) %>%
  summarize(mean_weight = mean(weight_value))

# Update length histograms to include mean lines
length_histogram_plot <- ggplot(length_samples_df, aes(x = length_value)) +
  geom_histogram(binwidth = 7, fill = "slategray2", color = "grey30", size = 0.2) +
  geom_vline(
    data = mean_length_values, 
    aes(xintercept = mean_length), 
    color = "salmon4", # Change to a similar color as your weight graph
    linetype = "dashed", 
    size = 0.8 # Adjust the line size to match your weight graph
  ) +
  labs(
    x = "Length (mm)", 
    y = "Frequency"
  ) +
  facet_wrap(~ month, ncol = 3) + # Use facet_wrap to arrange the plots by month
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold")
  )
length_histogram_plot

# Update weight histograms to include mean lines
weight_histogram_plot <- ggplot(weight_samples_df, aes(x = weight_value)) +
  geom_histogram(
    binwidth = 0.01, 
    fill = "darkseagreen3", 
    color = "black", 
    size = 0.2, # Adjust the border size of the histogram bars
    alpha = 0.7 # Adjust the transparency if needed
  ) +
  geom_vline(
    data = mean_weight_values, 
    aes(xintercept = mean_weight), 
    color = "salmon4", 
    linetype = "dashed", 
    size = 0.8 # Adjust the line size to match your length graph
  ) +
  labs(
    x = "Weight (kg)", 
    y = "Frequency"
  ) +
  facet_wrap(~ month, ncol = 3) + # Use facet_wrap to arrange the plots by month
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold")
  )
weight_histogram_plot

# Save plots
ggsave("Figures/length_histograms.pdf", plot = length_histogram_plot, width = 12, height = 12, dpi = 300)
ggsave("Figures/weight_histograms.pdf", plot = weight_histogram_plot, width = 12, height = 12, dpi = 300)

```

### Age-0 Count Function

```{r age0 count function, cache = TRUE}
calculate_total_salmon_consumed_multi_month_optimized <- function(season, samples_list, abund_func, cf_func, demo_func, prop, days_list, ind_cons_funcs, chinook_weight_funcs) {
  total_results <- list()  # List to store quantile results with change_name
  raw_samples <- list()    # List to store raw samples with change_name
  
  # Pre-generate the values that don't depend on the loop
  abund_values <- abund_func()
  cf_values <- cf_func()
  demo_values <- demo_func()
  prop_values <- prop
  
  # Pre-generate consumption values and Chinook weights for each month
  consumption_values <- sapply(ind_cons_funcs, function(func) func())
  chinook_weights <- sapply(chinook_weight_funcs, function(func) func())
  
  # Loop through each change in the adjusted samples
  for (change_name in names(samples_list)) {
    if (startsWith(change_name, season)) {
      # Extract the adjusted samples
      current_samples <- samples_list[[change_name]]
      
      # Calculate total salmon consumed over multiple months in one vectorized step
      total_consumed <- rowSums(
        sapply(1:length(days_list), function(i) {
          (abund_values * cf_values * demo_values * prop_values * days_list[i] * (consumption_values[i] / chinook_weights[i])) * current_samples
        })
      )
      
      # Store the quantile results with the change_name included
      quantiles <- quantile(total_consumed, probs = c(0.5, 0.025, 0.975, 0.25, 0.75))
      total_results[[change_name]] <- data.frame(change_name = change_name, 
                                                 median = quantiles[1], 
                                                 lower_95 = quantiles[2], 
                                                 upper_95 = quantiles[3], 
                                                 lower_25 = quantiles[4], 
                                                 upper_75 = quantiles[5])
      
      # Store the raw total consumed samples with the change_name included
      raw_samples[[change_name]] <- data.frame(change_name = change_name, total_consumed = total_consumed)
    }
  }
  
  # Convert the results to dataframes
  quantiles_df <- do.call(rbind, total_results)
  raw_samples_df <- do.call(rbind, raw_samples)
  
  return(list(quantiles = quantiles_df, raw_samples = raw_samples_df))
}
```

#### Winter

```{r winter demograpic count age 0, cache = TRUE}
# Winter calculations
# Winter calculations for Juvenile Females
count_WinterJuvF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "winter",
  samples_list = adjusted_samples_list,
  abund_func = WinAbund,
  cf_func = WinSprFallCF,
  demo_func = WinJuvDemo,
  prop = PropJuvF,
  days_list = c(31, 31, 28),  # December, January, February days
  ind_cons_funcs = list(IndDailyConsumptionJuvF, IndDailyConsumptionJuvF, IndDailyConsumptionJuvF),
  chinook_weight_funcs = list(DecJuvChinWt, JanJuvChinWt, FebJuvChinWt)
)

# Winter calculations for Juvenile Males
count_WinterJuvM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "winter",
  samples_list = adjusted_samples_list,
  abund_func = WinAbund,
  cf_func = WinSprFallCF,
  demo_func = WinJuvDemo,
  prop = PropJuvM,
  days_list = c(31, 31, 28),  # December, January, February days
  ind_cons_funcs = list(IndDailyConsumptionJuvM, IndDailyConsumptionJuvM, IndDailyConsumptionJuvM),
  chinook_weight_funcs = list(DecJuvChinWt, JanJuvChinWt, FebJuvChinWt)
)

# Winter calculations for Adult Females
count_WinterAdF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "winter",
  samples_list = adjusted_samples_list,
  abund_func = WinAbund,
  cf_func = WinSprFallCF,
  demo_func = WinFemDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # December, January, February days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(DecJuvChinWt, JanJuvChinWt, FebJuvChinWt)
)

# Winter calculations for Adult Males
count_WinterAdM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "winter",
  samples_list = adjusted_samples_list,
  abund_func = WinAbund,
  cf_func = WinSprFallCF,
  demo_func = WinMaleDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # December, January, February days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(DecJuvChinWt, JanJuvChinWt, FebJuvChinWt)
)

count_WinterAdF_df <- count_WinterAdF$raw_samples
count_WinterAdM_df <- count_WinterAdM$raw_samples
count_WinterJuvM_df <- count_WinterJuvM$raw_samples
count_WinterJuvF_df <- count_WinterJuvF$raw_samples
```

```{r total winter count age 0, cache=TRUE}
sum_consumed_by_change_group <- function(demographic_df_list) {
  # Ensure all dataframes are aligned by change_name
  for (i in seq_along(demographic_df_list)) {
    demographic_df_list[[i]] <- demographic_df_list[[i]] %>%
      arrange(change_name)
  }
  
  # Initialize an empty dataframe to store results
  result_df <- demographic_df_list[[1]]  # Start with the first dataframe's structure
  
  # Summing total_consumed across demographic groups for each change_name
  result_df$total_consumed <- 0  # Initialize the total_consumed column
  
  # Loop through each unique change_name and sum corresponding rows from each demographic dataframe
  unique_change_names <- unique(result_df$change_name)
  
  for (change in unique_change_names) {
    # Subset each dataframe for the current change_name
    filtered_dfs <- lapply(demographic_df_list, function(df) df[df$change_name == change, "total_consumed"])
    
    # Sum the total_consumed values across demographic groups for this change_name
    result_df[result_df$change_name == change, "total_consumed"] <- rowSums(do.call(cbind, filtered_dfs), na.rm = TRUE)
  }
  
  return(result_df)
}

# Example usage: summing across the winter demographic groups within each change_name
winter_demographic_count <- list(count_WinterAdF_df, count_WinterAdM_df, count_WinterJuvM_df, count_WinterJuvF_df)

# Get the summed results across demographic groups by change_name
merged_winter_samples <- sum_consumed_by_change_group(winter_demographic_count)

merged_winter_samples$change_name <- sub("^winter_", "", merged_winter_samples$change_name)

quantile_probs <- c(0.025, 0.25, 0.5, 0.75, 0.975)

# Calculate quantiles for each change group, including 25th and 75th percentiles
merged_winter_samples_sum <- merged_winter_samples %>%
  group_by(change_name) %>%
  summarize(
    ci_min = quantile(total_consumed, probs = 0.025, na.rm = TRUE),
    ci_low = quantile(total_consumed, probs = 0.25, na.rm = TRUE),
    avg = quantile(total_consumed, probs = 0.5, na.rm = TRUE),
    ci_high = quantile(total_consumed, probs = 0.75, na.rm = TRUE),
    ci_max = quantile(total_consumed, probs = 0.975, na.rm = TRUE)
  )

row.names(merged_winter_samples) <- NULL

```

```{r spring demographic count age0, cache = TRUE}
# Spring calculations for Juvenile Females
count_SpringJuvF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "spring",
  samples_list = adjusted_samples_list,
  abund_func = SprAbund,
  cf_func = WinSprFallCF,
  demo_func = SprJuvDemo,
  prop = PropJuvF,
  days_list = c(31, 31, 28),  # March, April, May days
  ind_cons_funcs = list(IndDailyConsumptionJuvF, IndDailyConsumptionJuvF, IndDailyConsumptionJuvF),
  chinook_weight_funcs = list(MarJuvChinWt, AprJuvChinWt, MayJuvChinWt)
)

# Spring calculations for Juvenile Males
count_SpringJuvM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "spring",
  samples_list = adjusted_samples_list,
  abund_func = SprAbund,
  cf_func = WinSprFallCF,
  demo_func = SprJuvDemo,
  prop = PropJuvM,
  days_list = c(31, 31, 28),  # March, April, May days
  ind_cons_funcs = list(IndDailyConsumptionJuvM, IndDailyConsumptionJuvM, IndDailyConsumptionJuvM),
  chinook_weight_funcs = list(MarJuvChinWt, AprJuvChinWt, MayJuvChinWt)
)

# Spring calculations for Adult Females
count_SpringAdF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "spring",
  samples_list = adjusted_samples_list,
  abund_func = SprAbund,
  cf_func = WinSprFallCF,
  demo_func = SprFemDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # March, April, May days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(MarJuvChinWt, AprJuvChinWt, MayJuvChinWt)
)

# Spring calculations for Adult Males
count_SpringAdM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "spring",
  samples_list = adjusted_samples_list,
  abund_func = SprAbund,
  cf_func = WinSprFallCF,
  demo_func = SprMaleDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # March, April, May days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(MarJuvChinWt, AprJuvChinWt, MayJuvChinWt)
)

count_SpringAdF_df <- count_SpringAdF$raw_samples
count_SpringAdM_df <- count_SpringAdM$raw_samples
count_SpringJuvM_df <- count_SpringJuvM$raw_samples
count_SpringJuvF_df <- count_SpringJuvF$raw_samples
```

```{r total spring count age 0, cache = TRUE}
# Assuming the spring dataframes are similar to the winter ones
spring_demographic_count <- list(count_SpringAdF_df, count_SpringAdM_df, count_SpringJuvM_df, count_SpringJuvF_df)

# Get the summed results across demographic groups by change_name for Spring
merged_spring_samples <- sum_consumed_by_change_group(spring_demographic_count)

# Remove the "spring_" prefix from change_name in the Spring dataframe
merged_spring_samples$change_name <- sub("^spring_", "", merged_spring_samples$change_name)

# Remove row names from Spring dataframe
row.names(merged_spring_samples) <- NULL

merged_spring_samples_sum <- merged_spring_samples %>%
  group_by(change_name) %>%
  summarize(
    ci_min = quantile(total_consumed, probs = 0.025, na.rm = TRUE),
    ci_low = quantile(total_consumed, probs = 0.25, na.rm = TRUE),
    avg = quantile(total_consumed, probs = 0.5, na.rm = TRUE),
    ci_high = quantile(total_consumed, probs = 0.75, na.rm = TRUE),
    ci_max = quantile(total_consumed, probs = 0.975, na.rm = TRUE)
  )

```

```{r summer demographic count age0, cache = TRUE}
# Summer calculations for Juvenile Females
count_SummerJuvF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "summer",
  samples_list = adjusted_samples_list,
  abund_func = SumAbund,
  cf_func = SummerCF,
  demo_func = SumJuvDemo,
  prop = PropJuvF,
  days_list = c(31, 31, 28),  # June, July, August days
  ind_cons_funcs = list(IndDailyConsumptionJuvF, IndDailyConsumptionJuvF, IndDailyConsumptionJuvF),
  chinook_weight_funcs = list(JunJuvChinWt, JulJuvChinWt, AugJuvChinWt)
)

# Summer calculations for Juvenile Males
count_SummerJuvM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "summer",
  samples_list = adjusted_samples_list,
  abund_func = SumAbund,
  cf_func = SummerCF,
  demo_func = SumJuvDemo,
  prop = PropJuvM,
  days_list = c(31, 31, 28),  # June, July, August days
  ind_cons_funcs = list(IndDailyConsumptionJuvM, IndDailyConsumptionJuvM, IndDailyConsumptionJuvM),
  chinook_weight_funcs = list(JunJuvChinWt, JulJuvChinWt, AugJuvChinWt)
)

# Summer calculations for Adult Females
count_SummerAdF <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "summer",
  samples_list = adjusted_samples_list,
  abund_func = SumAbund,
  cf_func = SummerCF,
  demo_func = SumFemDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # June, July, August days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(JunJuvChinWt, JulJuvChinWt, AugJuvChinWt)
)

# Summer calculations for Adult Males
count_SummerAdM <- calculate_total_salmon_consumed_multi_month_optimized(
  season = "summer",
  samples_list = adjusted_samples_list,
  abund_func = SumAbund,
  cf_func = SummerCF,
  demo_func = SumMaleDemo,
  prop = 1,  # Assuming full representation for adults
  days_list = c(31, 31, 28),  # June, July, August days
  ind_cons_funcs = list(IndDailyConsumptionAdF, IndDailyConsumptionAdF, IndDailyConsumptionAdF),
  chinook_weight_funcs = list(JunJuvChinWt, JulJuvChinWt, AugJuvChinWt)
)

count_SummerAdF_df <- count_SummerAdF$raw_samples
count_SummerAdM_df <- count_SummerAdM$raw_samples
count_SummerJuvM_df <- count_SummerJuvM$raw_samples
count_SummerJuvF_df <- count_SummerJuvF$raw_samples

```

```{r summer count age0, cache = TRUE}
summer_demographic_count <- list(count_SummerAdF_df, count_SummerAdM_df, count_SummerJuvM_df, count_SummerJuvF_df)

# Get the summed results across demographic groups by change_name for Summer
merged_summer_samples <- sum_consumed_by_change_group(summer_demographic_count)

merged_summer_samples <- as.data.frame(merged_summer_samples)
# Remove the "summer_" prefix from change_name in the Summer dataframe
merged_summer_samples$change_name <- sub("^summer_", "", merged_summer_samples$change_name)

merged_summer_samples_sum <- merged_summer_samples %>%
  group_by(change_name) %>%
  summarize(
    ci_min = quantile(total_consumed, probs = 0.025, na.rm = TRUE),
    ci_low = quantile(total_consumed, probs = 0.25, na.rm = TRUE),
    avg = quantile(total_consumed, probs = 0.5, na.rm = TRUE),
    ci_high = quantile(total_consumed, probs = 0.75, na.rm = TRUE),
    ci_max = quantile(total_consumed, probs = 0.975, na.rm = TRUE)
  )

# Remove row names from Summer dataframe
row.names(merged_summer_samples) <- NULL

# Print unique change_name values for Summer
```
### Seasonal sensitivity plot count of Age-0 Chinook Salmon 

```{r plot of age0 count by sensitivity changes, cache = TRUE}
merged_winter_samples_sum$Season <- "Winter"
merged_spring_samples_sum$Season <- "Spring"
merged_summer_samples_sum$Season <- "Summer"

age0_count_byseason <- rbind(merged_winter_samples_sum, merged_spring_samples_sum, merged_summer_samples_sum) 

age0_count_df <- age0_count_byseason %>%
  mutate(Change_Simplified = case_when(
    grepl("10%_Up", change_name) ~ "10% Increase",
    grepl("10%_Down", change_name) ~ "10% Decrease",
    grepl("25%_Up", change_name) ~ "25% Increase",
    grepl("25%_Down", change_name) ~ "25% Decrease",
    grepl("50%_Up", change_name) ~ "50% Increase",
    grepl("50%_Down", change_name) ~ "50% Decrease",
    grepl("Original_Proportion", change_name) ~ "Original Proportion"
  ))

age0_count_byseason_changes <- age0_count_df %>%
  filter(!grepl("Original Proportion", Change_Simplified))
age0_count_byseason_original <- age0_count_df %>%
  filter(grepl("Original Proportion", Change_Simplified))

change_plot_age0count <- ggplot(age0_count_byseason_changes, aes(x = Season, y = avg)) +
  geom_bar(stat = "identity", fill = "gray") +
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2, color = "black", linewidth = 0.8) +
  # Add smaller confidence interval (ci_low and ci_high) - draw this second to bring it to the front, set width to 0 to remove T's
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0, color = "blue", size = 6) +
  labs(title = "Count of Age-0 Chinook sensitivity test",
       x = "Season",
       y = "Total Count of Age-0 Chinook") +
  theme_minimal() +
  facet_wrap(~ Change_Simplified, scales = "fixed", ncol = 2)  # Set scales to "fixed" for a uniform y-axis
change_plot_age0count


# Plot only the original proportions
original_plot_age0count <- ggplot(age0_count_byseason_original, aes(x = Season, y = avg)) +
  geom_bar(stat = "identity", fill = "gray",color = "black") +
  # Add larger confidence interval (ci_min and ci_max)
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2, color = "black", size = 0.8) +
  # Add smaller confidence interval (ci_low and ci_high)
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0, color = "sienna3", size = 6) +
  labs(
       x = "Season",
       y = "Total Count of Age-0 Chinook") +
  theme_minimal()

original_plot_age0count

kable(age0_count_byseason_original)

ggsave("Figures/Figure 7.pdf", original_plot_age0count, units = "in", width = 8, height = 5, dpi = 300)

```

## Overall calculations

```{r combined age0 count data, cache = TRUE}
combined_df <- merged_winter_samples %>%
  mutate(season = "Winter") %>%
  bind_rows(
    merged_spring_samples %>% mutate(season = "Spring"),
    merged_summer_samples %>% mutate(season = "Summer")
  )

# Get unique change names
unique_change_names <- unique(combined_df$change_name)

# Initialize empty lists to store wide dataframes and summary statistics
wide_dfs <- list()
summarized_stats_list <- list()

# Loop through each unique change_name and perform operations
for (change in unique_change_names) {
  # Filter combined_df and pivot to wide format
  change_wide <- combined_df %>%
    filter(change_name == change) %>%
    pivot_wider(names_from = season, values_from = total_consumed, values_fn = list) %>%
    unnest(cols = c(Winter, Spring, Summer))
  
  # Calculate total consumed for the year
  change_wide$total_consumed_yearly <- rowSums(change_wide[, c("Winter", "Spring", "Summer")], na.rm = TRUE)
  
  # Store the wide dataframe
  wide_dfs[[change]] <- change_wide
  
  # Calculate summary statistics
  summarized_stats_list[[change]] <- quantile(
    change_wide$total_consumed_yearly, 
    probs = c(0.5, 0.025, 0.975, 0.25, 0.75), 
    na.rm = TRUE
  )
}

# Create a list of change groups to bind together
change_groups <- c("Change_10%_Up", "Change_10%_Down", "Change_25%_Up", "Change_25%_Down", 
                   "Change_50%_Up", "Change_50%_Down", "Original_Proportion")

# Bind wide dataframes for each change group with an additional change_name column
combined_wide_df <- bind_rows(lapply(change_groups, function(change) {
  wide_dfs[[change]] %>% mutate(change_name = change)
}))

# Bind summary statistics for each change group
overall_count_age0_summary <- bind_rows(lapply(change_groups, function(change) {
  data.frame(change_name = change, t(summarized_stats_list[[change]]))
}))
```

### bootstrapping

```{r sensitivity bootstrapping on age-0 count, cache = TRUE}
age0_count_output <- combined_wide_df %>% select(change_name, total_consumed_yearly)

# Function to perform bootstrapping
run_bootstrap_analysis <- function(original_dist, change_dist, n_sims = 2000) {
  # Calculate observed difference
   dist_1<- log(change_dist)
   dist_2<- log(original_dist)

  
  observed_diff <- mean(dist_1) - mean(dist_2)
  
  # Combine both distributions
  combined <- c(dist_1, dist_2)
  
  # Initialize a vector to store the resampled differences
  diffs <- numeric(n_sims)
  
  # Bootstrap resampling
  for (i in 1:n_sims) {
    shuffled <- sample(combined, length(combined), replace = FALSE)
    dist_1_resample <- shuffled[1:length(dist_1)]
    dist_2_resample <- shuffled[(length(dist_1) + 1):(length(combined))]
    diffs[i] <- mean(dist_1_resample) - mean(dist_2_resample)
  }
  
  # Calculate the p-value
  p_value <- mean(abs(diffs) >= abs(observed_diff))
  
  # Return observed difference and p-value
  return(c(observed_diff = observed_diff, p_value = p_value))
}

# Extract the original distribution for "Original_Proportion" from age0_count_output
original_dist <- age0_count_output %>%
  filter(change_name == "Original_Proportion") %>%
  pull(total_consumed_yearly)

# List of changes to compare against the original
changes_to_compare <- c("Change_10%_Up", "Change_10%_Down", "Change_25%_Up", 
                        "Change_25%_Down", "Change_50%_Up", "Change_50%_Down")

# Initialize an empty data frame to store results
age0_count_bootresults <- data.frame(Change_Comparison = character(), Observed_Difference = numeric(), P_Value = numeric(), stringsAsFactors = FALSE)

# Loop through each change and run the bootstrap analysis
for (change in changes_to_compare) {
  comparison_dist <- age0_count_output %>%
    filter(change_name == change) %>%
    pull(total_consumed_yearly)
  
  # Check if the comparison distribution is not empty
  if (length(comparison_dist) == 0) {
    cat("No data for", change, "\n")
  } else {
    # Run the bootstrap analysis
    result <- run_bootstrap_analysis(original_dist, comparison_dist)
    
    # Add the result to the results dataframe
    age0_count_bootresults <- rbind(age0_count_bootresults, data.frame(Change_Comparison = change,
                                         Observed_Difference = result['observed_diff'],
                                         P_Value = result['p_value']))
  }
}

# Print the results

overall_count_age0_summary_table <- overall_count_age0_summary %>%
  rename("Median" = "X50.", "ci_min" = "X2.5.", "ci_max" = "X97.5.") %>%
  mutate(
    Median = paste0(
      formatC(Median, format = "f", big.mark = ",", digits = 0), 
      " (", 
      formatC(ci_min, format = "f", big.mark = ",", digits = 0), 
      "-", 
      formatC(ci_max, format = "f", big.mark = ",", digits = 0), 
      ")"
    )
  ) %>%
  as.data.frame() %>% rename(c("Change_Comparison" = "change_name")) %>% select(Change_Comparison,Median)

kable(overall_count_age0_summary_table)

age0_count_bootresults$inverse_log_values <- exp(age0_count_bootresults$Observed_Difference)
age0_count_bootresults$percent_change <- (age0_count_bootresults$inverse_log_values-1)*100

age0_count_bootresults$base_age0avg <- overall_count_age0_summary %>%
  filter(change_name == "Original_Proportion") %>%
  pull('X50.')

age0_count_bootresults$expected_value <- age0_count_bootresults$base_age0avg * age0_count_bootresults$inverse_log_values

age0_count_bootresults_table <- merge(overall_count_age0_summary_table, age0_count_bootresults, by = "Change_Comparison", all = TRUE)

kable(age0_count_bootresults_table)
```

```{r count age-0 overall plots, cache = TRUE}

# Rename columns for better clarity
colnames(overall_count_age0_summary) <- c("Change", "50%", "2.5%", "97.5%", "25%", "75%")



# Select the median (50%) column for plotting
plot_data <- overall_count_age0_summary %>%
  select(Change, `50%`, `2.5%`, `97.5%`, `25%`, `75%`) 

plot_data$Change <- factor(
  plot_data$Change,
  levels = c("Original_Proportion", "Change_10%_Down", "Change_10%_Up", "Change_25%_Down", "Change_25%_Up","Change_50%_Down", "Change_50%_Up"),  # Specify the original order
  labels = c("Base Model", "10% Decrease", "10% Increase", "25% Decrease", "25% Increase","50% Decrease", "50% Increase")  # Specify the new labels
)

# Create the bar plot with error bars for confidence intervals
count_plot <-  ggplot(plot_data, aes(x = Change, y = `50%`, fill = Change)) +
  geom_bar(stat = "identity", color = "black", fill = "grey") +
  geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), width = 0.2, color = "black", size = 1) +
  geom_errorbar(aes(ymin = `25%`, ymax = `75%`), width = 0, color = "sienna3", size = 6) +
  labs(
       y = "Count of Age-0 Chinook consumed", 
       x = "Change Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
count_plot

ggsave("Figures/Figure9.pdf", plot = count_plot, width = 8, height = 5, dpi = 300)

```

# Daily consumption of Chinook salmon per individual Steller sea lion

Alongside looking at the overall consumption by Steller sea lions along the coast of Washington, we also calculated daily rates of the number of age-0 Chinook salmon consumed by an individual Steller sea lion depending on age, sex, and month.

## Daily biomass of age-0 Chinook salmon consumed per SSL demographic {.tabset}

```{r daily biomass consumed by demographic group, cache =TRUE}
# Helper function to calculate biomass or salmon count
calculate_consumption <- function(demo_function, chin_weight, age_og, n = 2000) {
  # Ensure chin_weight and age_og are scalars and applied element-wise across replicates
  replicates <- replicate(n, demo_function())  # Generate 10,000 replicates
  replicates_adjusted <- replicates / chin_weight * age_og  # Apply scalar adjustments
  quantile(replicates_adjusted, probs = c(0.5, 0.025, 0.975))  # Return quantiles
}

# Loop through demographics and months
demographics <- list(JuvF = IndDailyConsumptionJuvF, 
                     JuvM = IndDailyConsumptionJuvM, 
                     AdF = IndDailyConsumptionAdF, 
                     AdM = IndDailyConsumptionAdM)

months <- c("Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")
weights <- c(DecJuvChinWt(), JanJuvChinWt(), FebJuvChinWt(), MarJuvChinWt(), AprJuvChinWt(),
             MayJuvChinWt(), JunJuvChinWt(), JulJuvChinWt(), AugJuvChinWt())
age_og <- list(winter = winter_age0_og, spring = spring_age0_og, summer = summer_age0_og)

# Dataframe to store results
results_df <- data.frame()

for (demo_name in names(demographics)) {
  demo_func <- demographics[[demo_name]]
  
  for (i in seq_along(months)) {
    month <- months[i]
    weight <- weights[i]
    season <- ifelse(i <= 3, "winter", ifelse(i <= 6, "spring", "summer"))
    
    # Biomass calculation
    biomass <- calculate_consumption(demo_func, 1, age_og[[season]])
    
    # Salmon count calculation
    salmon_count <- calculate_consumption(demo_func, weight, age_og[[season]])
    
    # Store the results in a dataframe for both biomass and salmon count
    quantile_df_biomass <- data.frame(
      demo_month = paste(demo_name, month, sep = "_"),
      type = "biomass", 
      median = biomass[1],
      ci_lower = biomass[2],
      ci_upper = biomass[3]
    )
    
    quantile_df_salmon <- data.frame(
      demo_month = paste(demo_name, month, sep = "_"),
      type = "salmon_count", 
      median = salmon_count[1],
      ci_lower = salmon_count[2],
      ci_upper = salmon_count[3]
    )
    
    # Append both dataframes to the results_df
    results_df <- as.data.frame(rbind(results_df, quantile_df_biomass))
    results_df <- as.data.frame(rbind(results_df, quantile_df_salmon))
  }
}
```

### Summary Graph

```{r plot of daily biomass consumed by demographic group, cache=TRUE}

row.names(results_df) <- NULL
combined_salmon_count <- results_df %>% filter(type == "salmon_count") %>% select(-type)

combined_salmon_count <- separate(combined_salmon_count, col = demo_month, into = c( 'demo', 'Month'))

# Step 2: Combine `median`, `ci_lower`, and `ci_upper` into one column
combined_salmon_count_format <- combined_salmon_count %>%
  mutate(value = paste0(round(median, 1), " (", round(ci_lower, 1), "-", round(ci_upper, 1), ")")) %>%
  select(demo, Month, value)

# Step 3: Pivot wider to have `demographic` as columns and `month` as rows
combined_salmon_count_wide <- combined_salmon_count_format %>%
  pivot_wider(names_from = demo, values_from = value)

combined_salmon_count_wide %>%
  mutate(across(where(is.character), ~ format(., digits = 1))) %>%
  kable()



individual_dailysal_count_df_plot <- combined_salmon_count
individual_dailysal_count_df_plot$avg <- individual_dailysal_count_df_plot$median
individual_dailysal_count_df_plot$ci_min <- individual_dailysal_count_df_plot$ci_lower
individual_dailysal_count_df_plot$ci_max <- individual_dailysal_count_df_plot$ci_upper



individual_dailysal_count_df_plot$Month <- factor(individual_dailysal_count_df_plot$Month, levels =  c('Dec','Jan', 'Feb', 'Mar', 'Apr', 'May','Jun', 'Jul', 'Aug'), labels = c('December','January', 'February', 'March', 'April', 'May','June', 'July', 'August'))

write.csv(individual_dailysal_count_df_plot, file="code outputs/daily_count_consumed.csv")

p2 <- ggplot(data=individual_dailysal_count_df_plot, aes(x=Month, y=avg, ymin = ci_min, ymax = ci_max, group = demo, fill = demo))+
  geom_line(aes(colour = demo)) + 
  geom_point(stat="identity", aes(colour = demo)) +
  geom_ribbon(aes(ymin = ci_min, ymax = ci_max, color = demo), alpha=0.1, 
              linetype="dashed") + 
  expand_limits(y=0)  + 
  xlab('Month') + 
  ylab('Daily Age-0 Chinook Consumed per individual SSL')

p2
```
